---
title: "SF ridehail shift start location"
author: "Greg Macfarlane"
date: "9/24/2021"
output: html_document
---

We'll do this again, using the mlogit package for R.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mlogit)
library(modelsummary)
```

I got some data from Yufei
```{r loadata}
(d <- read_csv("data/estimation_data.zip"))
```

In a “long” dataset, there should be an ID column that identifies the choice
maker, and alternative column that identifies which options they have, and a
chosen column that identifies which alternative they chose.

```{r stats}
# number of choice makers
length(unique(d$ID))

d %>% group_by(ID) %>% 
  summarise(
    n = n(), # number of rows per person
    chosen = sum(chosen), # number of choices per person
    n_alts = length(unique(TAZ))
  )
```


This dataset is not well-specified. How is it possible for an individual
driver to have chosen four different starting locations? Is it because they have 
multiple shifts? Which variable identifies the shift number?

Let's take a guess that the `start` column could help us out.
```{r better-id}
d %>% group_by(ID, start) %>% 
  summarise(
    n = n(), # number of rows per person
    chosen = sum(chosen), # number of choices per person
    n_alts = length(unique(TAZ))
  )
```

Okay, now we are cooking. There is still a problem where some agents apparently have 
repeated alternatives. That's got to change; is there a compelling reason why 30 alternatives
is better than 10? 

Oh well, let's move forward. We'll remove the duplicate rows and make sure we keep
the choice.

```{r idx}
idx <- d %>% 
  filter(ID < 10) %>%
  mutate(ID = str_c(ID, start, sep = "-")) %>% 
  
  # keep only one row from the ID-TAZ pair, but make sure it's the chosen row
  group_by(ID, TAZ) %>%
  arrange(-chosen, .by_group = TRUE) %>%
  slice(1)  %>%
  
  dfidx(idx = c("ID", "TAZ"))
  
```

With that cleaning done, estimating a destination choice model is straightforward.
One thing to note is that you cannot have any alternative-specific coefficients 
(including the intercept). So we add a `-1` to the ASC specification.

```{r models}
mymodels <- list(
  "Base" = mlogit(chosen ~ POP + EMP | -1, data = idx),
  "Land Use" = mlogit(chosen ~ TOTALPARK + SFDU + MFDU | -1, data = idx),
  "Both" = mlogit(chosen ~ POP + EMP + TOTALPARK + SFDU + MFDU | -1, data = idx)
)
```

```{r results}
modelsummary(mymodels, estimate  = "{estimate} {stars}", statistic = "({statistic})",
             notes = "t-statistic in parentheses", fmt = "%.4f")
```

